The Effectiveness of Strategies to Contain SARS-CoV-2: Testing, Vaccinations, and NPIs

### 2022-03-11 src/paper/report.tex
Incorporate Janos' comments.
Final version for resubmission.

In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on age
\citep{Davies2020}.\label{r.2.b} A possible infection progresses as shown in
Figure~\ref{fig:model_disease_progression}. We differentiate between an initial period
of infection without being infectious or showing symptoms, being infectious
(presymptomatic or asymptomatic), showing symptoms, requiring intensive care, and
recovery or death \citep[similar to][]{Aleta2020}. The probabilities of transitioning
between these states depend on age; their duration is random and calibrated to medical
literature (for a detailed description see Supplementary
Material~\ref{sub:data_course_of_disease}). Conditional on the type of contact,
infectiousness is independent of age \citep{Jones2021}.

@@

We include two types of tests. Polymerase chain reaction (PCR) tests reveal whether an
individual is infected or not; there is no uncertainty to the result. PCR tests require
at least one day to be processed and there are aggregate capacity constraints. In
contrast, rapid antigen tests yield immediate results.
\label{r.1.a}\Copy{r.1.a}{Specificity and sensitivity of these tests is set according to
data analyzed in \citet{Bruemmer2021,Scheiblauer2021,Özcürümez2021}; sensitivity depends
on the timing of the test relative to the onset of infectiousness as in
\citet{Smith2021}. We provide a robustness to different assumptions in
Appendix~\ref{subsec:robustness_rapid_test_sensitivity}.} After a phase-in period, all
tests that are demanded will be performed. Figure~\ref{fig:model_rapid_tests} shows our
model for rapid test demand. Schools may require staff and students to be tested
regularly. Rapid tests may be offered by employers to on-site workers. Individuals may
demand tests for private reasons, which include having plans to meet other people,
showing symptoms of CoViD-19, and a household member having tested positively for the
virus. We endow each agent with an individual compliance parameter. This parameter
determines whether she takes up rapid tests.\footnote{Positive test results or symptoms
lead most individuals to reduce their contacts; this is why tests impact the actual
contacts in Figure~\ref{fig:model_description}.}

@@

A similar-sized effect---42\% in the decomposition---comes from rapid testing. Here, it
is crucial to differentiate between recorded cases and actual cases. Additional testing
means that additional infections will be recorded which would otherwise remain
undetected. Figure~\ref{fig:2021_scenarios_recorded} shows that this effect is large and
may persist for some time. Until late April, recorded cases are higher in the scenario
with rapid testing alone when compared to the setting where none of the three mechanisms
are turned on. The effect on total cases, however, is visible immediately in
Figure~\ref{fig:2021_scenarios_newly_infected}. Despite the fact that only 10\% of the
population performed weekly rapid tests in March on average, new infections on April~1
would have been reduced by 53\% relative to the scenario without vaccinations, rapid
tests, or seasonality. \label{r.1.b}\Copy{r.1.b}{In
    Appendix~\ref{subsec:robustness_rapid_test_sensitivity}, we provide a detailed analysis
    of whether our results are robust regarding the sensitivity parameters we assume for
    rapid tests. Even if we take a very pessimistic stance, the effect is only reduced from
    42\% to 38\%.}

>>>>

In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on age
\citep{Davies2020,Goldstein2021}.\label{r.2.b} A possible infection progresses as shown
in Figure~\ref{fig:model_disease_progression}. We differentiate between an initial
period of infection without being infectious or showing symptoms, being infectious
(presymptomatic or asymptomatic), showing symptoms, requiring intensive care, and
recovery or death \citep[similar to][]{Aleta2020}. The probabilities of transitioning
between these states depend on age; their duration is random and calibrated to medical
literature (for a detailed description see Supplementary
Material~\ref{sub:data_course_of_disease}). Conditional on the type of contact,
infectiousness is independent of age \citep{Jones2021}.

@@

We include two types of tests. Polymerase chain reaction (PCR) tests reveal whether an
individual is infected or not; there is no uncertainty to the result. PCR tests require
at least one day to be processed and there are aggregate capacity constraints. In
contrast, rapid antigen tests yield immediate results.
\label{r.1.a}\Copy{r.1.a}{Specificity and sensitivity of these tests is set according to
    data analyzed in \citet{Bruemmer2021,Scheiblauer2021,Özcürümez2021}; sensitivity depends
    on the timing of the test relative to the onset of infectiousness as in
    \citet{Smith2021}. We analyse robustness to different assumptions in
    Appendix~\ref{subsec:robustness_rapid_test_sensitivity}.} After a phase-in period, all
tests that are demanded will be performed. Figure~\ref{fig:model_rapid_tests} shows our
model for rapid test demand. Schools may require staff and students to be tested
regularly. Rapid tests may be offered by employers to on-site workers. Individuals may
demand tests for private reasons, which include having plans to meet other people,
showing symptoms of CoViD-19, and a household member having tested positively for the
virus. We endow each agent with an individual compliance parameter. This parameter
determines whether she takes up rapid tests.\footnote{Positive test results or symptoms
    lead most individuals to reduce their contacts; this is why tests impact the actual
    contacts in Figure~\ref{fig:model_description}.}

@@

A similar-sized effect---42\% in the decomposition---comes from rapid testing. Here, it
is crucial to differentiate between recorded cases and actual cases. Additional testing
means that additional infections will be recorded which would otherwise remain
undetected. Figure~\ref{fig:2021_scenarios_recorded} shows that this effect is large and
may persist for some time. Until late April, recorded cases are higher in the scenario
with rapid testing alone when compared to the setting where none of the three mechanisms
are turned on. The effect on total cases, however, is visible immediately in
Figure~\ref{fig:2021_scenarios_newly_infected}. Despite the fact that only 10\% of the
population performed weekly rapid tests in March on average, new infections on April~1
would have been reduced by 53\% relative to the scenario without vaccinations, rapid
tests, or seasonality. \label{r.1.b}\Copy{r.1.b}{In
    Appendix~\ref{subsec:robustness_rapid_test_sensitivity}, we provide a detailed analysis
    of whether our results are robust regarding the sensitivity parameters we assume for
    rapid tests. Even if we take a pessimistic stance, the effect is only reduced from 42\%
    to 38\%.}

### 2022-03-11 src/rev1_reply_comments.tex
Final version for resubmission.

\definecolor{darkblue}{rgb}{0, 0, 0.8}
\hypersetup{ colorlinks=true, linkcolor=darkblue, anchorcolor=darkblue,
    citecolor=darkblue, filecolor=darkblue, menucolor=darkblue, runcolor=darkblue,
    urlcolor=darkblue }
\usepackage{enumerate}
\usepackage{caption}
\usepackage{booktabs,caption,rotating}
\usepackage[capposition=top]{floatrow}
% \usepackage{bbm}
\usepackage{epstopdf}
\usepackage{longtable, tabularx, array}
\usepackage{colortbl}

@@

We thank the editors and referees for their excellent comments. Revising the
paper to address them has allowed us to substantially improve our paper.

% strongest change came through sensitivity comment
The largest adjustment was warrented with respect to the sensitivity of rapid
tests. We fully agree that our calibration appears optimistic in light of the
studies that have come out since our submission. We therefore verified that 
our results and conclusions are not affected when repeating the simulations 
with rapid tests with a much lower sensitivity. For that, we added a robustness
check with sensitivities that are derived from more recent
sources, relying on the progression of viral load over the course of an
infection and rapid test sensitivity with respect to different viral loads. 
These new sensitivities are on average significantly lower than those in our 
main specification. We are happy to report that even the most conservative 
estimates only lead to a small decrease in the estimated role of rapid tests.

The remaining comments have allowed us to improve our text significantly to be
clearer and better referenced. We also updated our citations for papers that we
had cited as pre-prints and that have been published in the meantime.
In our view the text has become clearer, more legible and better referenced 
as a result.

@@

@@

\response{%
    We would like to thank the reviewer for these insightful comments. First of all, we
    agree that the references we used have become outdated in the meantime. Note that we
    submitted the original manuscript in early September 2021; the analysis had been
    finalised in June of last year. Indeed, in the line of the additional evidence, the
    numbers we use for test sensitivity appear to be in the upper part of plausible
    values. We have followed your suggestion and report detailed robustness analyses in
    the paper and its Appendix.

@@

    \textit{\Paste{r.1.a}}

@@

    \textit{\Paste{r.1.b}}

@@

\response{We have added such a declaration at the end of the paper:\\[2ex]

@@

          \response{%
              Many thanks, it is of course true that vaccines do not stop transmission
              completely. We do not assume that, either. However, we have not been able
              to find data on the way in which this affects different population groups
              for the wild type and the $\alpha$-variant.

              In our model, we thus simplify the way vaccines reduce transmission risk.
              In particular, we assume that that 75\% of individuals achieve sterile
              immunity through a vaccination while the remaining 25\% remain fully
              susceptible. Firstly, the existing evidence supports the
              assumption that all four vaccines approved in the EU (Biontech/Pfizer,
              Moderna, AstraZeneca and Johnson \& Johnson) provide similar levels of
              protection in the first few months after vaccination against the wild type
              and the $\alpha$ variant (for details see Appendix~\ref{r.2.c}, starting
              on Page~\pageref{r.2.c}), which is the only time period and variant our
              simulations cover. Secondly, given the surprisingly good performance of
              rapid tests we tried to err on the side of being too optimistic with our
              modelling of vaccine performance. This is why we chose the perfect
              protection from vaccination for a large part of the population. Thirdly,
              reinfections are not a big issue for the time period that we look at. The
              incidences were small and the chance of more than one exposure to Covid
              are negligibly small. Thus, the value added by a probabilistic model of
              vaccination protection is very small for the period that we cover. }
\end{enumerate}
}

>>>>

\definecolor{darkblue}{rgb}{0, 0, 0.8}
\hypersetup{ colorlinks=true, linkcolor=darkblue, anchorcolor=darkblue,
    citecolor=darkblue, filecolor=darkblue, menucolor=darkblue, runcolor=darkblue,
    urlcolor=darkblue }
\usepackage{enumerate}
\usepackage{caption}
\usepackage{booktabs,caption,rotating}
\usepackage[capposition=top]{floatrow}
% \usepackage{bbm}
\usepackage{epstopdf}
\usepackage{longtable, tabularx, array}
\usepackage{colortbl}
\usepackage{placeins}

@@

We thank the editors and referees for their excellent comments. Revising the
paper to address them has allowed us to substantially improve our paper. \\[1ex]

The largest adjustment was warranted with respect to the sensitivity of rapid tests. We
fully agree that our calibration appears optimistic in light of the studies that have
come out since our submission. We therefore verified that our results and conclusions
are not affected when repeating the simulations under a much lower sensitivity of rapid
tests. For that, we added a robustness check with sensitivities that are derived from
more recent sources, relying on the progression of viral load over the course of an
infection and rapid test sensitivity with respect to different viral loads. These new
sensitivities are on average significantly lower than those in our main specification.
We are happy to report that even the most conservative estimates lead to only a minor
decrease in the estimated role of rapid tests. \\[1ex]

The remaining comments have allowed us to improve our text significantly to be clearer
and better referenced. We also updated our citations for papers that we cited as
pre-prints and that have been published in the meantime. In our view the text has become
clearer, more legible and better referenced as a result.\\[1ex]

We now address the comments by the Editors and Reviewers in turn.

@@

\clearpage

@@

\response{%
    We would like to thank the reviewer for these insightful comments. First of all, we
    agree that the references we used have become outdated in the meantime. Note that we
    submitted the original manuscript in early September 2021; the analysis had been
    finalised in June of last year. Indeed, in the light of the additional evidence, the
    numbers we use for test sensitivity appear to be in the upper part of plausible
    values. We have followed your suggestion and report detailed robustness analyses in
    the paper and its appendix.

@@

    \textit{\Paste{r.1.a}}\\[1ex]

@@

    \textit{\Paste{r.1.b}}\\[1ex]

@@

\response{We have added such a declaration at the end of the paper:

@@

          \response{%
              Many thanks, it is of course true that vaccines reduce transmissions, but
              do not stop them completely. We do not assume that, either. However, we
              have not been able to find data on the way in which this affects different
              population groups for the wild type and the $\alpha$-variant.

              In our model, we thus simplify the way vaccines reduce transmission risk.
              In particular, we assume that that 75\% of individuals achieve sterile
              immunity through a vaccination while the remaining 25\% remain fully
              susceptible. Firstly, the existing evidence supports the assumption that
              all four vaccines approved in the EU (Biontech/Pfizer, Moderna,
              Oxford/AstraZeneca and Johnson \& Johnson) provide similar levels of
              protection in the first few months after vaccination against the wild type
              and the $\alpha$ variant (for details see Appendix~\ref{r.2.c}, starting
              on Page~\pageref{r.2.c}), which is the only time period and variant our
              simulations cover. Secondly, given the surprisingly good performance of
              rapid tests we tried to err on the side of being too optimistic with our
              modelling of vaccine performance. This is why we chose the perfect
              protection from vaccination for a large part of the population. Thirdly,
              reinfections are not a big issue for the time period that we look at. The
              incidences were small and the chance of more than one exposure to Covid
              are negligibly small. Thus, the value added by a probabilistic model of
              vaccination protection is very small for the period that we cover. At the
              same time, it comes at prohibitive computational costs.}
\end{enumerate}
}
